### 2020-02-14 CalNotes.tex
some correction

Data-driven approaches to the derivation of effective stochastic coarse-grained dynamics have become a standard tool in different areas such as molecular dynamics or atmosphere and ocean science \cite{KKP15,CoP09,YMV19}. When fitting data to a low dimensional coarse-grained model, one is faced with the problem of model misspecification; indeed, the data, coming from the full dynamics, is compatible with the coarse-grained model only at the time scales at which the effective dynamics is valid. This is related to the issue of high-frequency data that arises in econometrics \cite{AiJ14,AMZ05,ZMA05,OSP10}.

@@

For simple models in molecular dynamics, the effect of model misspecification was studied in a series of papers \cite{PaS07,PPS09,PPS12,GaS17,GaS18} under the assumption of scale separation. In particular, for Brownian particles moving in two-scale potentials it was shown that, when fitting data from the full dynamics to the homogenized equation, the maximum likelihood estimator (MLE) is asymptotically biased \cite[Theorem 3.4]{PaS07}. To be more precise, in the large sample size limit the MLE converges to the coefficients of the unhomogenized equation, rather than to those of the homogenized one. The bias of the MLE can be eliminated by subsampling at an appropriate rate, which lies between the two characteristic time scales of the problem \cite[Theorems 3.5 and 3.6]{PaS07}. This is similar to the estimation of the integrated stochastic volatility in the presence of market micro-structure noise, where the data has to be subsampled at an appropriate rate \cite{AMZ05,ZMA05}. The correct subsampling rate can be in some instances rather extreme with respect to the frequency of the data and lead to get rid of more than $99\%$ of the data. As the intuition suggests, this increases significantly the bias of the estimator, which is usually taken care of with additional bias corrections and variance reduction procedures.

@@

We now briefly present the classical methodology for estimating the drift coefficient. Let $T > 0$ and let $X^\epl \defeq (X^\epl_t, 0\leq t \leq T)$ be a realization of the solution of \eqref{eq:SDE_MS} up to final time. Girsanov's change of measure formula applied to \eqref{eq:SDE_HOM} allows to write the likelihood of $X^\epl$ given a drift coefficient $A$ as
\begin{equation}\label{eq:Likelihood}
p(X^\epl \mid A) = \exp\left(-\frac{I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
I(X^\epl \mid A) = \int_0^T A \cdot V'(X^\epl_t) \dd X^\epl_t + \frac12 \int_0^T \left( A \cdot V'(X^\epl_t) \right)^2 \dd t.
\end{equation}
Minimizing the functional $I(X^\epl \mid A)$ with respect to $A$ therefore gives the maximum likelihood estimator (MLE) of $A$, which can be formally computed in closed form as
\begin{equation}\label{eq:MLE}
	\widehat A^\epl(T) \defeq \arg \min_{A \in \R^N} I(X^\epl \mid A) = M^{-1}h,
\end{equation}
where $M\in\R^{N\times N}$ and $h\in\R^N$ are defined as
\begin{equation}\label{eq:MandH}
M = \frac1{2\Sigma T} \int_0^T V'(X^\epl_t) \otimes V'(X^\epl_t) \dd t, \quad h = \frac1{2\Sigma T} \int_0^T V'(X^\epl_t) \dd X^\epl_t,
\end{equation}
and where $\otimes$ denotes the outer product in $\R^N$. Let us now state the assumptions which will be employed throughout the rest of our work. In particular, we consider the same dissipative setting as \cite[Assumption 3.1]{PaS07}.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ satisfy
	\begin{enumerate}
		\item $p \in \mathcal C^\infty(\R) \cap L^\infty(\R)$ and is $L$-periodic for some $L > 0$,
		\item\label{as:regularity_diss} $V_i \in \mathcal C^\infty(\R)$ for all $i=1, \ldots, N$ is polynomially bounded from above and bounded from below, and there exist $a,b > 0$ such that
		\begin{equation}
		-\sum_{i=1}^N V_i'(x)x \leq a - bx^2.
		\end{equation} 
		\item $V'$ is Lipschitz continuous on $\R$: there exists a constant $C > 0$ such that
		\begin{equation}
			\norm{V'(x) - V'(y)}_2 \leq C\abs{x - y},
		\end{equation} 
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and there exists $\bar \lambda > 0$ such that $\lambda_{\min}(M) \geq \bar \lambda$.
	\end{enumerate}
\end{assumption}

@@

\begin{theorem}\label{thm:Bias} Let Assumption \ref{as:regularity} hold and let $X^\epl_0$ be distributed according to the invariant measure of the process $X_t^\epl$ solution of \eqref{eq:SDE_MS}. Then
	\begin{equation}
		\lim_{\epl \to 0}\lim_{T \to \infty} \widehat A^\epl(T) = \alpha, \quad \text{a.s.},
	\end{equation}
	where $\alpha$ is the drift coefficient of equation \eqref{eq:SDE_MS}.
\end{theorem}

@@

\begin{example}\label{ex:OrnUhl} A closed form solution of \eqref{eq:FPsystem} can be obtained in a simple homogenized case with the dimension of the parameter $N=1$. Let $p(y) = 0$ and the parameters $\alpha$, $\sigma$ be replaced respectively by $A$ and $\Sigma$. Then, if $V(x) = x^2/2$, equation \eqref{eq:FPsystem} has the analytical solution 
\begin{equation}
\rho^0(x,z) = \frac{1}{C_{\rho^0}} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right),
\end{equation}
where
\begin{equation}
C_{\rho^0} = \int_{\R} \int_{\R} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right) \dd x \dd z.
\end{equation}
This is the density of a multivariate normal distribution $\mathcal N(0, \Gamma)$, where the covariance matrix is given by
\begin{equation}
\Gamma = \frac{\Sigma}{A (1 + A\delta)} \begin{pmatrix} 1+A\delta & 1 \\ 1 & 1 \end{pmatrix}.
\end{equation}
Let us remark that this distribution can be obtained from direct computations involving Gaussian processes. In particular, it is known that $X \sim \mathcal{GP}(\mu_t, \mathcal C(t, s))$, where at stationarity $\mu_t = 0$ and
\begin{equation}
	\mathcal C(t, s) = \frac{\Sigma}{A} e^{-A|t-s|}.
\end{equation}
The basic properties of Gaussian processes imply that $Z_t$ is a Gaussian process, and that the couple $(X_t, Z_t)^\top$ is a Gaussian process, too, whose mean and covariance are computable explicitly.
\end{example}

@@

Let us fix a Gaussian prior $\mu_0 = \mathcal N(A_0, C_0)$ on $A$, where $A_0 \in \R^N$ and $C_0 \in \R^{N\times N}$ is symmetric positive definite. Then, given a final time $T > 0$, the posterior distribution $\mu_{T,\epl}$ admits a density $p(A \mid X^\epl)$ with respect to the Lebesgue measure which satisfies
\begin{equation}
p(A \mid X^\epl) = \frac1{Z^\epl} \, p(X^\epl \mid A) \, p_0(A),
\end{equation}
where $Z^\epl$ is the normalization constant, $p_0$ is the density of $\mu_0$, and where the likelihood $p(X^\epl \mid A)$ is given in \eqref{eq:Likelihood}. The log-posterior density is therefore given by
\begin{equation}\label{eq:Posterior}
\log p(A \mid X^\epl) = -\log Z^\epl - T A \cdot h - \frac{T}{2} A \cdot M A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0),
\end{equation}
where $M$ and $h$ are defined in \eqref{eq:MandH}.
Since the log-posterior density is quadratic in $A$, the posterior is Gaussian, and it is therefore sufficient to determine its mean and covariance to fully characterize it. We denote by $m_{T,\epl}$ and $C_{T,\epl}$ the mean and covariance matrix, respectively. Completing the squares in the log-posterior density, we formally obtain
\begin{equation}
\begin{aligned}
C_{T,\epl}^{-1} &= C_0^{-1} + T M, \\
C_{T,\epl}^{-1}m_{T,\epl} &= C_0^{-1}A_0 - T h. 
\end{aligned}
\end{equation}
Under Assumption \ref{as:regularity}, one can show that the posterior at time $T > 0$ is indeed given by $\mu_{T,\epl} = \mathcal N(m_{T,\epl}, C_{T,\epl})$. We are now interested in the asymptotic limit of the posterior distribution for $T \to \infty$ and for $\epl \to 0$.

We can now state the main result for asymptotic convergence of the posterior distribution.
\begin{theorem}\label{thm:BayesianBias} Under Assumption \ref{as:regularity}, the posterior measure $\mu_{T,\epl} = \mathcal N(m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
		\lim_{\epl \to 0}\lim_{T \to \infty} \mu_{T,\epl}(\cdot) = \delta(\cdot - \alpha), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R^N$, where $\delta$ is the Dirac measure on $\R^N$.
\end{theorem}
\begin{proof} Let us first consider the covariance matrix. Hua's identity yields
	\begin{equation}
	C_{T,\epl} = T^{-1} \left(M^{-1} - Q^{-1}\right),
	\end{equation}
	where 
	\begin{equation}
	Q = M + T M C_0 M.
	\end{equation}
	Let us first remark that due to the hypothesis on $M$ and the ergodic theorem it holds for all $T > 0$
	\begin{equation}
		\norm{M^{-1}}_2 \leq \frac1{\bar \lambda}.
	\end{equation}
	We now have that for generic symmetric positive definite matrices $R$ and $S$ it holds
	\begin{equation}
		\norm{(R+S)^{-1}}_2 \leq \norm{S^{-1}}_2.
	\end{equation}
	Applying this inequality to $Q^{-1}$, we obtain
	\begin{equation}
		\norm{Q^{-1}}_2 \leq \frac1T \norm{(MC_0M)^{-1}}_2 \leq \frac1T \norm{M^{-1}}_2^2 \norm{C_0^{-1}}_2 = \frac1{T\bar \lambda^2} \norm{C_0^{-1}}_2,
	\end{equation}
	which implies
	\begin{equation}
		\lim_{T \to \infty}\norm{Q^{-1}}_2 = 0,
	\end{equation}
	and due to the triangle inequality
	\begin{equation}\label{eq:CovarianceShrink}
		\lim_{T \to \infty}\norm{C_{T,\epl}}_2 = 0.
	\end{equation}
	We proved that in the limit for $T \to \infty$ the covariance shrinks to zero independently of $\epl$. We now consider the mean. First, we remark that the triangle inequality yields
	\begin{equation}
		\norm{m_{T,\epl} - \alpha}_2 \leq \norm{m_{T,\epl} - \widehat A^\epl(T)}_2 + \norm{\widehat A^\epl(T) - \alpha}_2.
	\end{equation}
	For the second term, Theorem \ref{thm:Bias} implies 
	\begin{equation}
		\lim_{\epl \to 0} \lim_{T \to \infty}\norm{\widehat A^\epl(T) - \alpha}_2 = 0, \quad \text{a.s.}
	\end{equation}
	Let us now consider the first term.	Replacing the expression of the maximum likelihood estimator \eqref{eq:MLE} and due to the Cauchy--Schwarz and triangle inequalities, we obtain
	\begin{equation}
	\begin{aligned}
		\norm{m_{T,\epl} - \widehat A^\epl(T)}_2 &= \frac1T\norm{M^{-1}C_0^{-1}A_0 - Q^{-1}\left(C_0^{-1}A_0 - Th \right)}_2\\
		&\leq \frac1{T\bar \lambda} \norm{C_0^{-1}}_2 \left(\norm{A_0}_2 + \frac1{\bar \lambda}\norm{h}_2 + \frac1{T\bar \lambda} \norm{C_0^{-1}}_2\norm{A_0}_2\right).
	\end{aligned}
	\end{equation}
	Moreover, the ergodic theorem and the strong law of large numbers for martingales guarantee that $\norm{h}_2$ is bounded a.s. for $T \to \infty$. Therefore
	\begin{equation}
	\lim_{T\to\infty} \norm{m_{T,\epl} - \widehat A^\epl(T)}_2 = 0, \quad \text{a.s.},
	\end{equation}
	independently of $\epl$. Finally, 
	\begin{equation}
		\lim_{\epl \to 0} \lim_{T \to \infty}\norm{m_{T,\epl} - \alpha}_2 = 0, \quad \text{a.s.},
	\end{equation}
	which, together with \eqref{eq:CovarianceShrink}, implies the desired result. 
\end{proof}

@@

In this section, we present how to correct the faulty behaviour highlighted by Theorem \ref{thm:BayesianBias} with filtering techniques. In particular, we employ the same strategy that led to the estimator $\widetilde A^\epl_k(T)$ presented in Section \ref{sec:FilterMLE_2}. Therefore, we modify the likelihood function as
\begin{equation}
	\widetilde p(X^\epl \mid A) = \exp\left(-\frac{\widetilde I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
	\widetilde I(X^\epl \mid A) =\int_0^T A \cdot V'(Z^\epl_t) \dd X^\epl_t + \int_0^T \left(A \cdot V'(X^\epl_t)\right)^2 \dd t,
\end{equation}
so that the estimator $\widetilde A^\epl_k(T)$ introduced in \eqref{eq:AHatMixedTilde} is the MLE based on the new likelihood $\widetilde p(X^\epl \mid A)$. With this likelihood function, we obtain the modified posterior $\widetilde \mu_{T,\epl} = \mathcal N(\widetilde m_{T, \epl}, C_{T, \epl})$, whose parameters are given by
\begin{equation}
\begin{aligned}
	C_{T, \epl}^{-1} &= C_0^{-1} + T M, \\
	C_{T,\epl}^{-1}\widetilde m_{T,\epl} &= C_0^{-1}A_0 - T \widetilde h. 
\end{aligned}	
\end{equation}
Let us remark that the posterior $\widetilde \mu_{T,\epl}$ has the same covariance as $\mu_{T,\epl}$. The following result now holds.
\begin{theorem} Let the Assumptions of Theorem \ref{thm:mainTheoremTilde} hold. Then, the modified posterior measure $\widetilde\mu_{T,\epl} = \mathcal N(\widetilde m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
	\lim_{\epl \to 0}\lim_{T \to \infty} \widetilde \mu_{T,\epl}(\cdot) = \delta(\cdot - A), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R^N$, where $\delta$ is the Dirac measure on $\R^N$.
\end{theorem}
\begin{proof} The proof follows from the proof of Theorem \ref{thm:BayesianBias} and from Theorem \ref{thm:mainTheoremTilde}.
\end{proof}

>>>>

Data-driven approaches to the derivation of effective stochastic coarse-grained dynamics have become a standard tool in different areas such as molecular dynamics or atmosphere and ocean science \cite{KKP15,CoP09,YMV19}. When fitting data to a low dimensional coarse-grained model, one is faced with the problem of model misspecification; indeed, the data, coming from the full dynamics, are compatible with the coarse-grained model only at the time scales at which the effective dynamics is valid. This is related to the issue of high-frequency data that arises in econometrics \cite{AiJ14,AMZ05,ZMA05,OSP10}.

@@

For simple models in molecular dynamics, the effect of model misspecification was studied in a series of papers \cite{PaS07,PPS09,PPS12,GaS17,GaS18} under the assumption of scale separation. In particular, for Brownian particles moving in two-scale potentials it was shown that, when fitting data from the full dynamics to the homogenized equation, the maximum likelihood estimator (MLE) is asymptotically biased \cite[Theorem 3.4]{PaS07}. To be more precise, in the large sample size limit the MLE converges to the coefficients of the unhomogenized equation, rather than to those of the homogenized one. The bias of the MLE can be eliminated by subsampling at an appropriate rate, which lies between the two characteristic time scales of the problem \cite[Theorems 3.5 and 3.6]{PaS07}. This is similar to the estimation of the integrated stochastic volatility in the presence of market micro-structure noise, where the data have to be subsampled at an appropriate rate \cite{AMZ05,ZMA05}. The correct subsampling rate can be in some instances rather extreme with respect to the frequency of the data and lead to get rid of more than $99\%$ of the data. As the intuition suggests, this increases significantly the bias of the estimator, which is usually taken care of with additional bias corrections and variance reduction procedures.

@@

We now briefly present the classical methodology for estimating the drift coefficient. Let $T > 0$ and let $X^\epl \defeq (X^\epl_t, 0\leq t \leq T)$ be a realization of the solution of \eqref{eq:SDE_MS} up to final time. Girsanov's change of measure formula applied to \eqref{eq:SDE_HOM} allows to write the likelihood of $X^\epl$ given a drift coefficient $A$ as
\begin{equation}\label{eq:Likelihood}
p(X^\epl \mid A) = \exp\left(-\frac{I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
I(X^\epl \mid A) = \int_0^T A \cdot V'(X^\epl_t) \dd X^\epl_t + \frac12 \int_0^T \left( A \cdot V'(X^\epl_t) \right)^2 \dd t.
\end{equation}
Minimizing the functional $I(X^\epl \mid A)$ with respect to $A$ therefore gives the maximum likelihood estimator (MLE) of $A$, which can be formally computed in closed form as
\begin{equation}\label{eq:MLE}
	\widehat A^\epl(T) \defeq \arg \min_{A \in \R^N} I(X^\epl \mid A) = M^{-1}h,
\end{equation}
where $M\in\R^{N\times N}$ and $h\in\R^N$ are defined as
\begin{equation}\label{eq:MandH}
M = \frac1T \int_0^T V'(X^\epl_t) \otimes V'(X^\epl_t) \dd t, \quad h = \frac1T \int_0^T V'(X^\epl_t) \dd X^\epl_t,
\end{equation}
and where $\otimes$ denotes the outer product in $\R^N$. Let us now state the assumptions which will be employed throughout the rest of our work. In particular, we consider the same dissipative setting as \cite[Assumption 3.1]{PaS07}.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ satisfy
	\begin{enumerate}
		\item $p \in \mathcal C^\infty(\R) \cap L^\infty(\R)$ and is $L$-periodic for some $L > 0$,
		\item\label{as:regularity_diss} $V_i \in \mathcal C^\infty(\R)$ for all $i=1, \ldots, N$ is polynomially bounded from above and bounded from below, and there exist $a,b > 0$ such that
		\begin{equation}
		-\sum_{i=1}^N V_i'(x)x \leq a - bx^2.
		\end{equation} 
		\item $V'$ is Lipschitz continuous, i.e. there exists a constant $C > 0$ such that
		\begin{equation}
			\norm{V'(x) - V'(y)}_2 \leq C\abs{x - y},
		\end{equation} 
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and there exists $\bar \lambda > 0$ such that $\lambda_{\min}(M) \geq \bar \lambda$.
	\end{enumerate}
\end{assumption}

@@

\begin{theorem}\label{thm:Bias} Let Assumption \ref{as:regularity} hold and let $X^\epl_0$ be distributed according to the invariant measure of the process $X^\epl$ solution of \eqref{eq:SDE_MS}. Then
	\begin{equation}
		\lim_{\epl \to 0}\lim_{T \to \infty} \widehat A^\epl(T) = \alpha, \quad \text{a.s.},
	\end{equation}
	where $\alpha$ is the drift coefficient of equation \eqref{eq:SDE_MS}.
\end{theorem}

@@

\begin{example}\label{ex:OrnUhl} A closed form solution of \eqref{eq:FPsystem} can be obtained in a simple homogenized case with the dimension of the parameter $N=1$. Let $p(y) = 0$ and the parameters $\alpha$, $\sigma$ be replaced respectively by $A$ and $\Sigma$. Then, if $V(x) = x^2/2$, equation \eqref{eq:FPsystem} has the analytical solution 
\begin{equation}
\rho^0(x,z) = \frac{1}{C_{\rho^0}} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right),
\end{equation}
where
\begin{equation}
C_{\rho^0} = \int_{\R} \int_{\R} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right) \dd x \dd z.
\end{equation}
This is the density of a multivariate normal distribution $\mathcal N(0, \Gamma)$, where the covariance matrix is given by
\begin{equation}
\Gamma = \frac{\Sigma}{A (1 + A\delta)} \begin{pmatrix} 1+A\delta & 1 \\ 1 & 1 \end{pmatrix}.
\end{equation}
Let us remark that this distribution can be obtained from direct computations involving Gaussian processes. In particular, it is known that $X \sim \mathcal{GP}(\mu_t, \mathcal C(t, s))$, where at stationarity $\mu_t = 0$ and
\begin{equation}
	\mathcal C(t, s) = \frac{\Sigma}{A} e^{-A|t-s|}.
\end{equation}
The basic properties of Gaussian processes imply that $Z$ is a Gaussian process, and that the couple $(X, Z)^\top$ is a Gaussian process, too, whose mean and covariance are computable explicitly.
\end{example}

@@

Let us fix a Gaussian prior $\mu_0 = \mathcal N(A_0, C_0)$ on $A$, where $A_0 \in \R^N$ and $C_0 \in \R^{N\times N}$ is symmetric positive definite. Then, given a final time $T > 0$, the posterior distribution $\mu_{T,\epl}$ admits a density $p(A \mid X^\epl)$ with respect to the Lebesgue measure which satisfies
\begin{equation}
p(A \mid X^\epl) = \frac1{Z^\epl} \, p(X^\epl \mid A) \, p_0(A),
\end{equation}
where $Z^\epl$ is the normalization constant, $p_0$ is the density of $\mu_0$, and where the likelihood $p(X^\epl \mid A)$ is given in \eqref{eq:Likelihood}. The log-posterior density is therefore given by
\begin{equation}\label{eq:Posterior}
\log p(A \mid X^\epl) = -\log Z^\epl - \frac{T}{2\Sigma} A \cdot h - \frac{T}{4\Sigma} A \cdot M A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0),
\end{equation}
where $M$ and $h$ are defined in \eqref{eq:MandH}.
Since the log-posterior density is quadratic in $A$, the posterior is Gaussian, and it is therefore sufficient to determine its mean and covariance to fully characterize it. We denote by $m_{T,\epl}$ and $C_{T,\epl}$ the mean and covariance matrix, respectively. Completing the squares in the log-posterior density, we formally obtain
\begin{equation}
\begin{aligned}
C_{T,\epl}^{-1} &= C_0^{-1} + \frac{T}{2\Sigma} M, \\
C_{T,\epl}^{-1}m_{T,\epl} &= C_0^{-1}A_0 - \frac{T}{2\Sigma} h. 
\end{aligned}
\end{equation}
Under Assumption \ref{as:regularity}, one can show that the posterior at time $T > 0$ is indeed given by $\mu_{T,\epl} = \mathcal N(m_{T,\epl}, C_{T,\epl})$. We are now interested in the asymptotic limit of the posterior distribution for $T \to \infty$ and for $\epl \to 0$.

We can now state the main result for asymptotic convergence of the posterior distribution.
\begin{theorem}\label{thm:BayesianBias} Under Assumption \ref{as:regularity}, the posterior measure $\mu_{T,\epl} = \mathcal N(m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
		\lim_{\epl \to 0}\lim_{T \to \infty} \mu_{T,\epl}(\cdot) = \delta(\cdot - \alpha), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R^N$, where $\delta$ is the Dirac measure on $\R^N$.
\end{theorem}
\begin{proof} Let us first consider the covariance matrix. Hua's identity yields
	\begin{equation}
	C_{T,\epl} = \frac{2\Sigma}{T} \left(M^{-1} - Q^{-1}\right),
	\end{equation}
	where 
	\begin{equation}
	Q = M + \frac{T}{2\Sigma} M C_0 M.
	\end{equation}
	Let us first remark that due to the hypothesis on $M$ and the ergodic theorem it holds for all $T > 0$
	\begin{equation}
		\norm{M^{-1}}_2 \leq \frac1{\bar \lambda}.
	\end{equation}
	We now have that for generic symmetric positive definite matrices $R$ and $S$ it holds
	\begin{equation}
		\norm{(R+S)^{-1}}_2 \leq \norm{S^{-1}}_2.
	\end{equation}
	Applying this inequality to $Q^{-1}$, we obtain
	\begin{equation}
		\norm{Q^{-1}}_2 \leq \frac{2\Sigma}T \norm{(MC_0M)^{-1}}_2 \leq \frac{2\Sigma}T \norm{M^{-1}}_2^2 \norm{C_0^{-1}}_2 = \frac{2\Sigma}{T\bar \lambda^2} \norm{C_0^{-1}}_2,
	\end{equation}
	which implies
	\begin{equation}
		\lim_{T \to \infty}\norm{Q^{-1}}_2 = 0,
	\end{equation}
	and due to the triangle inequality
	\begin{equation}\label{eq:CovarianceShrink}
		\lim_{T \to \infty}\norm{C_{T,\epl}}_2 = 0.
	\end{equation}
	We proved that in the limit for $T \to \infty$ the covariance shrinks to zero independently of $\epl$. We now consider the mean. First, we remark that the triangle inequality yields
	\begin{equation}
		\norm{m_{T,\epl} - \alpha}_2 \leq \norm{m_{T,\epl} - \widehat A^\epl(T)}_2 + \norm{\widehat A^\epl(T) - \alpha}_2.
	\end{equation}
	For the second term, Theorem \ref{thm:Bias} implies 
	\begin{equation}
		\lim_{\epl \to 0} \lim_{T \to \infty}\norm{\widehat A^\epl(T) - \alpha}_2 = 0, \quad \text{a.s.}
	\end{equation}
	Let us now consider the first term.	Replacing the expression of the maximum likelihood estimator \eqref{eq:MLE} and due to the Cauchy--Schwarz and triangle inequalities, we obtain
	\begin{equation}
	\begin{aligned}
		\norm{m_{T,\epl} - \widehat A^\epl(T)}_2 &= \frac{2\Sigma}T\norm{M^{-1}C_0^{-1}A_0 - Q^{-1}\left(C_0^{-1}A_0 - \frac{T}{2\Sigma} h \right)}_2\\
		&\leq \frac{2\Sigma}{T\bar \lambda} \norm{C_0^{-1}}_2 \left(\norm{A_0}_2 + \frac1{\bar \lambda}\norm{h}_2 + \frac{2\Sigma}{T\bar \lambda} \norm{C_0^{-1}}_2\norm{A_0}_2\right).
	\end{aligned}
	\end{equation}
	Moreover, the ergodic theorem and the strong law of large numbers for martingales guarantee that $\norm{h}_2$ is bounded a.s. for $T \to \infty$. Therefore
	\begin{equation}
	\lim_{T\to\infty} \norm{m_{T,\epl} - \widehat A^\epl(T)}_2 = 0, \quad \text{a.s.},
	\end{equation}
	independently of $\epl$. Finally, 
	\begin{equation}
		\lim_{\epl \to 0} \lim_{T \to \infty}\norm{m_{T,\epl} - \alpha}_2 = 0, \quad \text{a.s.},
	\end{equation}
	which, together with \eqref{eq:CovarianceShrink}, implies the desired result. 
\end{proof}

@@

In this section, we present how to correct the faulty behaviour highlighted by Theorem \ref{thm:BayesianBias} with filtering techniques. In particular, we employ the same strategy that led to the estimator $\widetilde A^\epl_k(T)$ presented in Section \ref{sec:FilterMLE_2}. Therefore, we modify the likelihood function as
\begin{equation}
	\widetilde p(X^\epl \mid A) = \exp\left(-\frac{\widetilde I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
	\widetilde I(X^\epl \mid A) =\int_0^T A \cdot V'(Z^\epl_t) \dd X^\epl_t + \int_0^T \left(A \cdot V'(X^\epl_t)\right)^2 \dd t,
\end{equation}
so that the estimator $\widetilde A^\epl_k(T)$ introduced in \eqref{eq:AHatMixedTilde} is the MLE based on the new likelihood $\widetilde p(X^\epl \mid A)$. With this likelihood function, we obtain the modified posterior $\widetilde \mu_{T,\epl} = \mathcal N(\widetilde m_{T, \epl}, C_{T, \epl})$, whose parameters are given by
\begin{equation}
\begin{aligned}
	C_{T, \epl}^{-1} &= C_0^{-1} + \frac{T}{2\Sigma} M, \\
	C_{T,\epl}^{-1}\widetilde m_{T,\epl} &= C_0^{-1}A_0 - \frac{T}{2\Sigma} \widetilde h. 
\end{aligned}	
\end{equation}
Let us remark that the posterior $\widetilde \mu_{T,\epl}$ has the same covariance as $\mu_{T,\epl}$. The following result now holds.
\begin{theorem} Let the Assumptions of Theorem \ref{thm:mainTheoremTilde} hold. Then, the modified posterior measure $\widetilde\mu_{T,\epl} = \mathcal N(\widetilde m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
	\lim_{\epl \to 0}\lim_{T \to \infty} \widetilde \mu_{T,\epl}(\cdot) = \delta(\cdot - A), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R^N$, where $\delta$ is the Dirac measure on $\R^N$.
\end{theorem}
\begin{proof} The proof follows from the proof of Theorem \ref{thm:BayesianBias} and from Theorem \ref{thm:mainTheoremTilde}.
\end{proof}
